---
title: "p8105_hw3_qra2000"
author: Quinn Anderson
output: html_document
date: "2023-10-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
```

## Problem 1

```{r read in data}
devtools::install_github("p8105/p8105.datasets")
force=TRUE

library(tidyverse)
library(p8105.datasets)

data("instacart")
```

The instacart dataset is an anonymized dataset that includes data from more than 200,000 Instacart users. The dataset has `r nrow(instacart)` observations of `r ncol(instacart)` variables, where each row is a product that Instacart users can order. Key variables in this data include identifiers (`order_id`, `product_id`, `user_id`, `aisle_id`, and `department_id`) and information about each product in each order (`add_to_cart_order`, `reordered`, `order_number`, `order_dow`, `order_hour_of_day`, and `days_since_prior_order`.) For example, user id 112108 ordered Bulgarian Yogurt (product id 49302) at 10am (`order_hour_of_day`) on Wednesday (4th day of the week, `order_dow`), which user ID 112108 has ordered at least once prior. This item is one out of 8 (`add_to_cart_order`) included in `order_id` . 

There are 134 aisles in the instacart dataset. The "fresh vegetables" aisle has the most frequent orders, at 150,609 times ordered from.  

```{r number of aisles in dataset}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

OR?

```{r}
instacart |> 
  group_by(aisle) |> 
  summarize(n_obs = n()) |> 
  mutate(
    aisle_rank = min_rank((desc(aisle)))
  )
```

The plot below shows the number of items ordered in each aisle, limited to aisles with more than 10,000 items ordered in this dataset. 

```{r plot items ordered in each aisle}
instacart |>  
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  labs(
    title = "Number of items ordered by aisle",
    x = "Aisle",
    y = "Number of times ordered") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The table below shows the 3 most popular items in each of the aisles: "baking ingredients", "dog food care", and "packaged vegetable fruits", and the number of times each item is ordered. 

```{r}
instacart |> 
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")
  ) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(
    rank = min_rank((desc(n)))
  ) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |> 
  knitr::kable()
```

The table below shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. 

```{r}
instacart |> 
  filter(
    product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")
  ) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable()
```

## Problem 2

```{r}
data("brfss_smart2010")
```





