---
title: "p8105_hw3_qra2000"
author: Quinn Anderson
output: github_document
date: '`r format(Sys.time(), "%Y-%m-%d")`'
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%")

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

options(knitr.table.format = 'markdown')
```

## Problem 1

### Read in data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

### Dataset descriptions and questions

The instacart dataset is an anonymized dataset that includes data from more than 200,000 Instacart users. The dataset has `r nrow(instacart)` observations of `r ncol(instacart)` variables, where each row is a product that Instacart users can order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. There are also several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

The table below summarizes the number of items ordered from the aisles. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far. 

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

The plot below shows the number of items ordered in each aisle, limited to aisles with more than 10,000 items ordered in this dataset. 

```{r}
instacart |>  
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  labs(
    title = "Number of items ordered by aisle",
    x = "Aisle",
    y = "Number of times ordered") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The table below shows the 3 most popular items in each of the aisles: `baking ingredients`, `dog food care`, and `packaged vegetable fruits`, and the number of times each item is ordered. 

```{r}
instacart |> 
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank((desc(n)))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |> 
  knitr::kable()
```

The table below shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. The table has been formatted for human readers. Pink lady apples are typically ordered earlier in the day than Coffee Ice Cream (~11am) with the exception of day 5. 

```{r}
instacart |> 
  filter(
    product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour) |> 
  knitr::kable(digits = 2)
```

## Problem 2

### Read in data

```{r}
data("brfss_smart2010")
```

### Data cleaning

```{r}
brfss_smart2010_clean =
  brfss_smart2010 |> 
  janitor::clean_names() |>
  filter(topic == "Overall Health",
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) |> 
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE)) |>
  rename("state" = "locationabbr","county" = "locationdesc")
```

### Data discussion

```{r}
brfss_smart2010_clean |> 
  filter(year == "2002") |> 
  group_by(year, state) |> 
  summarize(nloc = n_distinct(geo_location)) |> 
  filter(nloc >= 7) |>
  arrange(desc(nloc)) |> 
  knitr::kable()

brfss_smart2010_clean |> 
  filter(year == "2010") |> 
  group_by(year, state) |> 
  summarize(nloc = n_distinct(geo_location)) |> 
  filter(nloc >= 7) |> 
  arrange(desc(nloc)) |> 
  knitr::kable()
```

In 2002, there were 6 states were observed at 7 or more locations: PA, MA, NJ, CT, FL, NC. 

In 2010, 14 states were observed at 7 or more locations: FL, NJ, TX, CA, MD, NC, NE, WA, MA, NY, OH, CO, PA, SC. 

Construct a dataset that is limited to `Excellent` responses, and contains, `year`, `state`, and a variable that averages the `data_value` across locations within a state. Make a "spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
bfrss_smart2010_excellent =
  brfss_smart2010_clean |> 
  filter(response == "Excellent") |> 
  select(year, state, county, data_value) |> 
  group_by(year, state) |> 
  summarize(mean_dv = mean(data_value))

bfrss_smart2010_excellent |>
  ggplot(aes(x = year, y = mean_dv, group = state)) + 
  geom_line(aes(color = state)) + 
  labs(
    title = "Average 'Excellent' Responses over Time per State",
    x = "Time (years)",
    y = "Average data_value",
    color = "States") +
  theme(plot.title = element_text(hjust = 0.5))
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_smart2010_clean |> 
  filter(year %in% c("2006", "2010", state == "NY")) |> 
  select(year, state, county, response, data_value) |> 
  ggplot(aes(x = response, y = data_value, color = response)) +
  geom_boxplot() +
  facet_grid(.~ year) +
  labs(
    title = "Distribution of data_value in NY State",
    x = "Response",
    y = "data_value") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Problem 3

### Read in, clean, tidy, and merge datasets

Read in and clean NHANES demographic data

```{r}
nhanes_covar = 
  read_csv("./data/nhanes_covar.csv", skip = 4)

nhanes_covar_clean =
  nhanes_covar |> 
  janitor::clean_names() |> 
  mutate(
    sex = recode(sex, `1` = "male", `2` = "female"),
    education = recode(education, `1` = "less than high school", `2` = "high school equivalent", `3` = "more than high school")) |> 
  drop_na(bmi, education) |> 
  filter(age > 21)
```

Read and clean NHANES accelerometer data.

```{r}
nhanes_accel = 
  read_csv("./data/nhanes_accel.csv")

nhanes_accel_clean = nhanes_accel |> 
  janitor::clean_names() |> 
  drop_na()
```

Combine both datasets.

```{r}
nhanes_data = merge(nhanes_covar_clean, nhanes_accel_clean)
```

### Data table and visualization, preliminary exploration

Produce reader friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category.

```{r}
nhanes_gender_education =
  nhanes_data |> 
  group_by(sex, education) |> 
  summarize(count = n()) |> 
  pivot_wider(
    names_from = "sex",
    values_from = "count") |> 
  knitr::kable()

ggplot(nhanes_data, aes(x = education, y = age, fill = sex)) + 
  geom_boxplot() +
  facet_grid(. ~ sex) +
  labs(
    title = "Age Distributions for Men and Women by Education",
    x = "Sex",
    y = "Age",
    color = "Sex") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 60, hjust = 1))
```

The plot above shows the age distribution by education category among the 250 total male and female participants in the NHANES study. The plot above shows the majority of the younger participants in the study across both male and females have an education more than high school, although slightly more female participants have education beyond high school.

### Data analyses

#### Total activity over the day

Using tidied dataset, aggregate across minutes to create "total activity" variable for each participant. Plot activities (y-axis) against age (x-axis). 

```{r}
nhanes_data |> 
  group_by(education, sex) |> 
  mutate(total_activity = rowSums(across(min1:min1440))) |> 
  ggplot(aes(x = age, y = total_activity, color = education)) +
  geom_point(alpha = .5) +
  facet_wrap(~education) +
  geom_smooth()+
  labs(
    title = "Total Activities against Age",
    x = "Age",
    y = "Activity") +
  theme(plot.title = element_text(hjust = 0.5))
```

The plot above shows the total activities compared to participant age from data in the NHANES study, separated by education level. This plot shows that total activity decreases as age increases across all three education levels. Participants wtih education less than high school have the highest total activities at the youngest age (20), then sharply decreases to the lowest total activities at the highest age (80). For participants with an education of high school equivalent, the highest total activity is among participants around 40 years old. For participants with an education beyond high school, the total activity plateaus for ages 20-40, then declines after age 60. 

#### Inspection activity over the course of the day

Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. 

```{r}
nhanes_data |> 
  pivot_longer(
    min1:min1440,
    names_to = "time", 
    values_to = "activity") |> 
  ggplot(aes(x = time, y = activity, color = sex)) +
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(.~ education) +
  labs(
    title = "24-hour activity time course for each education level across sex",
    x = "Time (minutes)",
    y = "Activity") +
  theme(plot.title = element_text(hjust = 0.5))
```

This plot shows the 24-hour activity time courses for male and female participants of the NHANES study across education level. The plot shows that male participants with education more than high school have the highest 24-total activity, followed by female participants with education more than high school. From this we can conclude the 

Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences. 







